A SAT-Based Counterexample Guided Method for Unbounded Synthesis

### 2016-01-29 paper.tex
Edited SAT formula explanation

    \begin{algorithmic}
        \algrestore{b1}
        \Function{findCandidate}{$p, s, k, T$}
            \State $f \gets $ \IfElse{$p = \texttt{cont}$}{\Call{treeFormula}{$T$}}{\Call{\textoverline{treeFormula}}{$T$}} \EndIfElse
            \State $sol \gets $ \Call{SAT}{$s(X_T) \land f$}
            \If{$sol = \texttt{unsat}$} 
                \If{\texttt{unbounded}} \Comment{Active only in the unbounded solver}
                    \State $\sigma \gets $ \Call{generalise}{$s$} \Comment{Expand $s$ to a set of states}
                    \State \IfElse{$p = \texttt{cont}$}{\Call{learn}{$\sigma, T$}}{\Call{\textoverline{learn}}{$\sigma, T$}} \EndIfElse
                \EndIf
                \State \Return $\emptyset$ \Comment{No candidate exists}
            \Else
                \State \Return $\{ \langle n, c \rangle | n \in $ \Call{nodes}{$T$} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
            \EndIf
        \EndFunction
        \algstore{b2}
    \end{algorithmic}

@@

Candidates are discovered by passing a formulation of the abstract game tree to
a SAT solver in $\textsc{findCandidate}$. This formula contains CNF encodings
of all of the unrolled runs represented by the tree and the winning condition
of the current player.  Runs are encoded by copying the transition relation for
every step in the abstract game. When playing for the controller, the SAT
solver searches for a satisfying assignment to the unfixed label variables in
tree so that none of the runs reaches the error state. The environment
formulation is satisfiable if any run does reach the error state.  The formula
is constructed recursively from the root of a tree by $\textsc{treeFormula}$
(see Algorithm~\ref{alg:treeFormula}).

Since the game tree formulation is passed to a SAT solver, both controllable
and uncontrollable unfixed labels will be existentially quantified. This means
that the SAT solver will find any way to win the game while both players are
cooperating. If no winning run exists in an abstract game even when the players
are cooperating then there is no winning run when the opponent is playing
adversarily. When a winning run is found, the actions chosen by the SAT solver
are used to refine the game tree. This is advantageous for many synthesis
problems where the game must be formalised as adversarial for correctness but
the final implementation will cooperate with its environment in the real world.
An example of such a system is a device driver that cooperates with the device
and OS to provide the interface between the two.

@@

When executing the unbounded solver, lines 18 and 19 become active in the
bounded solver. These lines call the learning procedures when the solver fails
to find a candidate for an abstract game tree. The states symbolically
represented by nodes in the tree are losing for whichever player could not find
a winning candidate and can be extracted from the tree using interpolation.

@@

\subsection{Choosing opponent moves}

One aspect of the algorithm proposed by Narodytska et al.~\cite{narodytska2014}
is that it mimics real-world uses of synthesis by allowing cooperation between
system and environment. Unfortunately, allowing the opponents to pick moves for
one another can frequently backfire. It is common when modelling real-world
systems to abstract over out of scope failures with transitions that
immediately determine the outcome of the game. For example, in a network driver
a request to send a packet may fail due to a disconnected wire, which might be
modelled as an environment controlled transition to a system-winning state.

This slows down the synthesis process as all erroneous transition are
accumulated into the abstract game tree. To mitigate this effect we have a
heuristic to guess non-erroneous labels as \emph{default moves} that are used
instead of allowing the opponent to choose. This does not effect the
correctness of the algorithm. If line~(28) previously returned \texttt{sat} and
is \texttt{unsat} with default moves then those moves would eventually have
been added to the abstraction in a refinement step. If the formula is
\texttt{unsat} even with the opponent choosing moves then it must be
\texttt{unsat} with default moves.

>>>>

    \begin{algorithmic}
        \algrestore{b1}
        \Function{findCandidate}{$p, s, k, T$}
        \State $\hat{T} \gets $ \Call{extend}{$T$} \Comment{Extend the tree with unfixed actions}
            \State $f \gets $ \IfElse{$p = \texttt{cont}$}{\Call{treeFormula}{$\hat{T}$}}{\Call{\textoverline{treeFormula}}{$\hat{T}$}} \EndIfElse
            \State $sol \gets $ \Call{SAT}{$s(X_{\hat{T}}) \land f$}
            \If{$sol = \texttt{unsat}$} 
                \If{\texttt{unbounded}} \Comment{Active only in the unbounded solver}
                    \State $\sigma \gets $ \Call{generalise}{$s$} \Comment{Expand $s$ to a set of states}
                    \State \IfElse{$p = \texttt{cont}$}{\Call{learn}{$\sigma, \hat{T}$}}{\Call{\textoverline{learn}}{$\sigma, \hat{T}$}} \EndIfElse
                \EndIf
                \State \Return $\emptyset$ \Comment{No candidate exists}
            \Else
                \State \Return $\{ \langle n, c \rangle | n \in $ \Call{nodes}{$T$} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
            \EndIf
        \EndFunction
        \algstore{b2}
    \end{algorithmic}

@@

The CEGAR loop depends on the ability to guess candidate partial strategies in
\textsc{findCandidate}. For this purpose we use the heuristic that a partial
strategy may be winning if for each \textsc{outcome} in the tree there exists a
play of controller and environment actions that is winning for the current
player. We can formulate this heuristic as a SAT query, which is constructed
recursively by $\textsc{treeFormula}$ ($\textsc{\textoverline{treeFormula}}$)
in Algorithm~\ref{alg:treeFormula}. The tree is first extended to the maximum
bound with \emph{unfixed} actions that allow existential choice by the SAT
solver. Then each game round in the tree is encoded in CNF by making copies of
the transition relation along with the winning condition of the current player.

The heuristic is based on the observation that for many synthesis problems in
the real world the final implementation will cooperate with its environment.
The existential search of the SAT solver simulates cooperation between the two
players.

%%%Candidates are discovered by passing a formulation of the abstract game tree to
%%%a SAT solver in $\textsc{findCandidate}$. This formula contains CNF encodings
%%%of all of the unrolled runs represented by the tree and the winning condition
%%%of the current player.  Runs are encoded by copying the transition relation for
%%%every step in the abstract game. When playing for the controller, the SAT
%%%solver searches for a satisfying assignment to the unfixed label variables in
%%%tree so that none of the runs reaches the error state. The environment
%%%formulation is satisfiable if any run does reach the error state.  The formula
%%%is constructed recursively from the root of a tree by $\textsc{treeFormula}$
%%%(see Algorithm~\ref{alg:treeFormula}).

%%%Since the game tree formulation is passed to a SAT solver, both controllable
%%%and uncontrollable unfixed labels will be existentially quantified. This means
%%%that the SAT solver will find any way to win the game while both players are
%%%cooperating. If no winning run exists in an abstract game even when the players
%%%are cooperating then there is no winning run when the opponent is playing
%%%adversarily. When a winning run is found, the actions chosen by the SAT solver
%%%are used to refine the game tree. This is advantageous for many synthesis
%%%problems where the game must be formalised as adversarial for correctness but
%%%the final implementation will cooperate with its environment in the real world.
%%%An example of such a system is a device driver that cooperates with the device
%%%and OS to provide the interface between the two.

@@

When executing the unbounded solver, lines 19 and 20 become active in the
bounded solver. These lines call the learning procedures when the solver fails
to find a candidate for an abstract game tree. The states symbolically
represented by nodes in the tree are losing for whichever player could not find
a winning candidate and can be extracted from the tree using interpolation.

@@

%%%\subsection{Choosing opponent moves}

%%%One aspect of the algorithm proposed by Narodytska et al.~\cite{narodytska2014}
%%%is that it mimics real-world uses of synthesis by allowing cooperation between
%%%system and environment. Unfortunately, allowing the opponents to pick moves for
%%%one another can frequently backfire. It is common when modelling real-world
%%%systems to abstract over out of scope failures with transitions that
%%%immediately determine the outcome of the game. For example, in a network driver
%%%a request to send a packet may fail due to a disconnected wire, which might be
%%%modelled as an environment controlled transition to a system-winning state.

%%%This slows down the synthesis process as all erroneous transition are
%%%accumulated into the abstract game tree. To mitigate this effect we have a
%%%heuristic to guess non-erroneous labels as \emph{default moves} that are used
%%%instead of allowing the opponent to choose. This does not effect the
%%%correctness of the algorithm. If line~(28) previously returned \texttt{sat} and
%%%is \texttt{unsat} with default moves then those moves would eventually have
%%%been added to the abstraction in a refinement step. If the formula is
%%%\texttt{unsat} even with the opponent choosing moves then it must be
%%%\texttt{unsat} with default moves.
